---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r}
#loading the data
mydata <- read.csv('C:\\Users\\Shengbing\\Documents\\R\\Reproducible_research\\assignment1\\RepData_PeerAssessment1\\activity\\activity.csv')
#remove missing values
#mydata <- mydata[complete.cases(mydata),]


```



## What is mean total number of steps taken per day?

**Both histogram and bargraph are shown below.**

```{r}
total.perday <- tapply(mydata$steps, mydata$date, sum, na.rm = T) #sum of steps per day
hist(total.perday, col = 'blue', breaks=15, main = 'Total steps each day (histogram)')
barplot(total.perday, col = 'blue', main = 'Total steps each day (barplot)')
mean.eachday <- tapply(mydata$steps, mydata$date, mean, na.rm = T)
mean.eachday
hist(mean.eachday, col = 'blue', main = 'mean steps each day')
median.eachday <- tapply(mydata$steps, mydata$date, median, na.rm = T) #there are a lot of #zeros each day, so the median is zero
median.eachday
hist(median.eachday, col = 'blue', breaks = 10, main = 'median steps each day')

```


## What is the average daily activity pattern?
```{r}
steps.each.interval <- tapply(mydata$steps, mydata$interval, mean, na.rm = T)
plot(names(steps.each.interval), steps.each.interval, type = 'l', col = 'blue', main = 'average daily activity pattern', xlab = '5-min-interval', ylab = 'average number of steps')
which.max(steps.each.interval)

```

## Imputing missing values
**Impute missing data using mean of each 5-minute interval**

*Values of total number of steps, mean and median total number of steps taken per day are different from the first part of the assignment. Imputing missing data removes a lot of zeros of the total daily number of steps.*

```{r}
#Calculate and report the total number of missing values in the dataset (i.e. the #total number of rows with NAs)
nrow(mydata) - sum(complete.cases(mydata))
#fill in all of the missing values in the dataset.
imputed.data = mydata
interval.mean = tapply(imputed.data$steps, imputed.data$interval, mean, na.rm = T)
for (i in names(interval.mean)) {
        imputed.data$steps[imputed.data$interval == as.integer(i) & is.na(imputed.data$steps)]= interval.mean[i]
}
#total, mean and median of imputed data
total.perday <- tapply(imputed.data$steps, imputed.data$date, sum) #sum of steps per day
hist(total.perday, col = 'red', breaks =15, main = 'Total steps each day (imputed)')
mean.eachday <- tapply(imputed.data$steps, imputed.data$date, mean)
mean.eachday
hist(mean.eachday, col = 'red', main = 'mean steps each day (imputed)')
median.eachday <- tapply(imputed.data$steps, imputed.data$date, median) 
median.eachday  #there are still a lot of zeros
hist(median.eachday, col = 'red', main = 'median steps each day (imputed)')
```


## Are there differences in activity patterns between weekdays and weekends?
**The answer is yes**
*It seems that in weekdays there is a major peak between 500 and 1000 minuts, but in weekend the activity is more spread*

```{r}
#install timeData package if necessary and load it
library(timeDate)
#Create a new factor variable in the dataset with two levels - "weekday" and #"weekend" 
imputed.data$isweekdays <- factor(isWeekday(as.Date(imputed.data$date), wday = 1:5), labels = c('weekend', 'weekday'))
#use ggplot2 to make a panel plot
library(ggplot2)
ggplot(data = imputed.data, aes(x = interval, y = steps)) + stat_summary(col = 'blue', fun.y = mean, geom = 'line') + facet_grid(isweekdays~.)

```
